### 2020-05-29 presentations/satRday2020/satRday2020.Rmd
some more changes

---
title: "dittodb"
subtitle: "making testing databases easy"
author: "Jonathan Keane · Socure"
institute: "Chicago satRdays 2020"
date: "2020-05-30 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", "Azuro.css"]
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      slideNumberFormat: ""
---
background-image: url("images/IMG_5793.png")
background-position: 50% 50%
background-size: cover

class: inverse, center, top

# What is testing?

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(emojifont) 
library(xaringanthemer)
style_mono_accent(
  base_color = "#1c5253",
  header_font_family = "Azuro",
  text_font_family = "Azuro",
  code_font_google   = google_font("Fira Mono"),
  text_font_size = "1.25rem",
  colors = c(white = "#FFFFFF", gray = "#898C9C"),
  extra_css = list(
    ".large" = list(
      "font-size" = "2.75rem"
    ),
    ".inverse > h1 > code" = list(
      "color" = "var(--inverse-header-color)"
    ),
    ".colophon" = list(
      "font-size" = "0.9rem",
      "position" = "absolute",
      "bottom" = 0
    ),
    ".flex-box" = list(
    "display" = "flex",
    "flex-direction" = "column",
    "justify-content" = "space-between",
    "height" = "70%"
    ),
    # the first p of .flext-box should have no top-margin
    ".flex-box > p:first-child" = list(
      "margin-top" = 0
    ),
    ".flex-box > p:last-of-type" = list(
      "padding-bottom" = "1em"
    ),
    ".seventyfive" = list(
      "width" = "75%",
      "margin" = "0 auto"
    ),
    ".title-slide h1" = list(
      "font-size" = "4rem",
      "margin-bottom" = "0rem"
      ),
    ".title-slide h2" = list(
      "font-size" = "1.75rem",
      "margin-top" = "0rem",
      "margin-bottom" = "4rem"
      ),
    ".title-slide h3" = list(
      "font-size" = "1.75rem",
      "margin-block-start" = "0rem",
      "margin-block-end" = "0rem"
      ),
    ".title-slide > h3:last-of-type" = list(
      "font-size" = "1rem",
      "margin-block-start" = "4rem",
      "margin-block-end" = "2rem"
    )
  )
)
```

---
class: inverse, center, middle

## A way to check that the code you wrote is doing what you expect it to.

---
# Easier said than done

<br /><br />
Any test is better than no tests

--
<br /><br />

but a good test is worth its (metaphorical) weight in gold

---
# A good test is ...
* quick to run
* expressive
* doesn't require special set up

--

# But databases are ...
* (relatively) slow
* rely on external state or set up
* requires a connection or custom setup

---
# CI + DBs = ☠️

Continuous integration (like GitHub Actions, Travis, AppVeyor) is wonderful, 
you can run your code on someone else's ☁️ 💻 
<br/> <br/>

--

But then you're running your code in an environment you have to set up.
--
 That you can't interactively iterate on.
<br/> <br/>

--

Even installing R dependencies can be a pain.

---
background-image: url("images/db-test-ci-db-hard-more-again.png")
background-position: 50% 50%
background-size: contain
class: center, bottom

---
background-image: url("images/db-test-ci-db-hard-more.png")
background-position: 0% 20%
background-size: contain
class: center, bottom

--

# I've slogged through this <br/> so you don't have to

---
background-image: url("images/_DSC1176.png")
background-position: 50% 50%
background-size: cover
class: center, top

# How does `dittodb` help?

---
# Imagine we had a function

```r
#' Get random airline(s)
#' ...
get_an_airline <- function(n = 1) {
  con <- DBI::dbConnect(RPostgres::Postgres(), dbname = "nycflights")
  on.exit(DBI::dbDisconnect(con))

  query <- paste0(
    "SELECT carrier, name FROM airlines ",
    "ORDER BY random() LIMIT ", 
    n
  )

  out <- dbGetQuery(con, query)

  if (out$name == "Testy McAirline") {
    stop("This is the test airline 🙃")
  }

  return(out)
}
```
---
# How can we test it?
Locally is “easy” we:
 * setup a local postgres server
 * make sure the correct users exist
 * add the test data
 * profit

--

On CI we “just” do the same:  
 * setup a local postgres server
 * make sure the correct users exist
 * add the test data
--
<br/>
<div class = "large">❌ ❌ ❌ ❌ ☠️ ☠️ ☠️ ❌ ❌ ❌ ❌</div>

---
# `dittodb` lets you ...

Write tests that act like they are connected to a real-live database using 
`with_mock_db({...})` without ever having to set one up.

<br/ >

```r
with_mock_db({
  test_that("We get one airline", {
    one_airline <- get_an_airline()
    expect_is(one_airline, "data.frame")
    expect_equal(nrow(one_airline), 1)
    expect_equal(one_airline$carrier, "9E")
    expect_equal(one_airline$name, "Endeavor Air Inc.")
  })
})
```

---

# Testing terminology aside

* mocks

* fixtures

* stubs

* spies

* fakes

* `r emoji("scream")`

???

* mocks -- objects pre-programmed with expectations which form a specification 
* fixtures -- data used by mocked objects to configure them or provide info 
* stubs -- provide canned answers to calls made during the test, usually not responding at all to anything outside what's programmed in for the test.
* spies -- stubs, but keep and provide info about how and if they were called
* fakes -- objects actually have working implementations, but usually take some shortcut which makes them not suitable for production (an in memory database is a good example)

Much of the above is from https://www.martinfowler.com/articles/mocksArentStubs.html

---

# Testing terminology aside

* mocks

* fixtures

.gray[
* stubs

* spies

* fakes

* `r emoji("scream")`
]

---
background-image: url("images/_DSC1098.png")
background-position: 50% 50%
background-size: cover
class: center, top

# Demo

???

If doing the demo, jump to slide 30

---
# Capturing DB interactions

Capture database interactions so that writing and editing the above is easier.

```r
capture_db_requests({
  one_airline <- get_an_airline()
})
```

<br />

Where `one_airline` is

```r
  carrier                 name
1      AS Alaska Airlines Inc.
```

---
# Capturing DB interactions

This response is saved to `testthat/tests/nycflights` in a file called `SELECT-c3427d.R`:

```r
structure(
  list(carrier = "AS", name = "Alaska Airlines Inc."), 
  class = "data.frame", 
  row.names = c(NA, -1L)
)
```

--

The path is `testthat/tests` and then the db name `nycflights` being connected to.

--

The filename starts with the SQL verb `SELECT`, followed by a hash that is unique to the query.

???

Because the hash is unique per-query that means this particular query will always return the same thing each time. But there are ways around this, for example by using a different mock paths.

---
# `dittodb` lets you ...
.flex-box[
So you can write your test like this

```r
with_mock_db({
  test_that("We get one airline", {
    one_airline <- get_an_airline()
    expect_is(one_airline, "data.frame")
    expect_equal(nrow(one_airline), 1)
    expect_equal(one_airline$carrier, "9E")
    expect_equal(one_airline$name, "Endeavor Air Inc.")
  })
})
```

.pull-right[... and it will run anywhere!]
]

---
# `dittodb` lets you ...

Write tests that include edge cases that “shouldn't” “happen” in your databse.

```r
with_mock_path(path = "error_db_fixt/", {
  with_mock_db({
    test_that("We get one airline", {
      expect_error(
        get_an_airline(),
        "This is the test airline \U1F643
      )
    })
  })
})
```
---
background-image: url("images/_DSC1098.png")
background-position: 50% 50%
background-size: cover
class: center, top

# It really is as easy as that!

---
# Mandatory plug
.flex-box[
## Socure is hiring — [socure.com/about/careers](https://www.socure.com/about/careers)

.center.seventyfive[![](https://www.socure.com/hubfs/Assets_2018/SOC-logo-color.svg)]
]

---

# Thanks!
.flex-box[
`dittodb` lives at [_dittodb.jonkeane.com_](https://dittodb.jonkeane.com)

Install `dittodb` with `remotes::install_github("jonkeane/dittodb")`

Extra special thanks to Mauricio Vargas @pachamaltese and thanks to [`httptest`](https://enpiar.com/r/httptest) for inspiration.

@jonkeane 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;
[_jonkeane.com_](https://jonkeane.com) 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;
[github.com/jonkeane](https://github.com/jonkeane)

.colophon[
created with the R package [**xaringan**](https://github.com/yihui/xaringan) · all photos taken by Jonathan Keane, all rights reserved <br />
presentation slides + content (except for photos) are licensed under [CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/) <br />
]
]